---
title: "Comparing prostate risk stratification tools"
date: '`r strftime(Sys.time(), "%A, %b %d, %Y")`'
output:
  rmarkdown::html_vignette:
    df_print: kable
    standalone: true
  pdf_document: 
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.width = 10, fig.height = 7)

```

```{css global-css, echo=FALSE}
img {
  margin: 0;
  padding: 0;
  max-width: 100%;
}
```

```{r library, include=FALSE}
library(drake)
library(tidyverse)
```

```{r}
loadd(risk_ncdb)
loadd(risk_seer)
loadd(seer_c_index)
loadd(ncdb_c_index)
```

Jordan H. Creed
Christelle 
Travis A. Gerke 

Department of Cancer Epidemiology, H. Lee Moffitt Cancer Center & Research Institute, 12902 USF Magnolia Drive, Tampa, FL, 33612, USA

Key Words - prostate; risk-startification

Corresponding author:
Travis A. Gerke, ScD
Department of Cancer Epidemiology
H. Lee Moffitt Cancer Center and Research Institute 
Tampa, FL 33612-9416
Phone: 
Fax:
Email: travis.gerke@moffitt.org

\newpage

**Abstract**

**Background:**

**Methods:**

**Results:**

**Conclusions:**

\newpage

**Introduction**

Pretreatment risk stratification tools are used to predict prostate cancer 
outcomes in men. However these tools were developed in cohorts with predominantly
European ancestry. Meanwhile African American men (AAM) are disproportionally affected
by prostate cancer, with AAM twice as likely than European American men (EAM) to
die from the disease [@seer]. 

The D'Amico risk stratification was developed in 1998 and is based on known 
prognostic factors and a literature review and classifies individuals into low,
intermediate and high risk of PSA failure [@damico1998]. 
The National Institute of Health and Care Excellence (NICE) created their own 
risk stratification for providing treatment recommendations to men with 
localised prostate cancer [@graham2014]. The European Association of Urology 
(EAU) guidelines aim to classify patients according to risk for biochemical 
recurrence following treatment [@mottet2017]. The GU Radiation Oncologists of
Canada (GUROC) also agree upon a low, intermediate and high risk stratification 
for localised prostate cancer [@lukka2001], while the American Urological 
Association (AUA) ...
The National Comprehensive Care Network (NCCN) recommends managing treatment
options and clinical decision making based off their risk stratification, which 
includes additional very high and very low risk classifications as well as 
further stratifying intermediate into favorable and unfavorable [@mohler2016].
The Cambridge Prognostic Groups (CPG) propose a five-level risk startification system 
for predicting prostate cancer specific mortality in non-metastatic prostate 
cancer cases [@gnanapragasam2016]. The Cancer of the Prostate Risk Assessment 
(CAPRA) score is a numeric risk score (ranging from one to ten) that assesses
biochemical recurrence specifically after radical prostatectomy [@cooperberg2005].
These risk stratification systems are all informed in some way by an individuals
PSA, Gleason, and stage. 

We used Surveillance, Epidemiology, and End Results (SEER) and National Cancer
Database (NCDB) prostate cancer databases to compare the performance of the 
following pretreatment risk stratification tools: D'Amico, NICE, EAU, GUROC, AUA,
NCCN, CPG and CAPRA, both overall and within European and African American men. 

**Paticipants and Methods**

*SEER*

SEER is a population-based database developed by the National Cancer Institute,
US Center for Disease Control and Prevention and regional and state cancer 
registeries and includes cancer cases from 18 states [@doll2018]. 

*NCDB*

NCDB is the result of the joint efforts of the American College of Surgeons and
the American Cancer Society and covers almost 70% of incident cancer cases
[@boffa2017]. 

**Statistical Analysis**

Risk classification was calculated for each risk stratification tool. For 
participants with T stage greater or lesser than those used in the original 
cutpoints, the nearest original point was used. Risk classifications were
determined using the values available. Cox proportional hazard models
(HR) and 95% confidence intervals for overall survival were conducted for each
risk tool. In addition to crude model, models were adjusted for age (continuous),
year of diagnosis (continuous), Charlson Comorbidity Index (continuous; not available for 
SEER), marital status (married/domestic partners vs single/separated/divorced/widowed;
not available for NCDB), education level (quartiles of adults who did not graduate high]
school; not available for SEER) and primary treatment (). Risk classification was assessed 
by the concordance index within each model, both overall and by year post diagnosis. 

Analyses were performed overall and stratified by self-reported race as well as by primary treatment.

As a sensitivity analysis, we performed the above-described analyses in prostate-specific 
deaths in SEER (data not available in NCDB).

All analyses were performed using R (version 3.5.0). 

**Results**

*SEER*

A total of `r length(unique(risk_seer$PUBCSNUM))` men are included in the 
prostate cancer data for SEER. The men are predominantly white (n =
`r nrow(risk_seer[risk_seer$RAC_RECA==1,])`, 
`r round(nrow(risk_seer[risk_seer$RAC_RECA==1,])/nrow(risk_seer)*100, 2)` %),
and during followup, `r nrow(risk_seer[risk_seer$os==1,])` 
(`r round(nrow(risk_seer[risk_seer$os==1,])/nrow(risk_seer)*100, 2)` %
) men died. Across all risk strifications, the overall crudec-index ranges from 
`r round(min(unlist(seer_c_index$concord_crude)), 3)` to 
`r round(max(unlist(seer_c_index$concord_crude)), 3)`. 
The lowest c-index was obsered for D'Amico, NICE and EAU at 0.604, while the 
highest was GCP at 0.693. 

*NCDB*

In NCDB, there were `r length(unique(risk_ncdb$PUF_CASE_ID))` particiapnts with 
prostate cancer, of which `r nrow(risk_ncdb[risk_ncdb$RACE==1,])` 
(`r round(nrow(risk_ncdb[risk_ncdb$os==1,])/nrow(risk_ncdb)*100, 2)` %) were 
white and `r nrow(risk_ncdb[risk_ncdb$os==1,])` 
(`r round(nrow(risk_ncdb[risk_ncdb$os==1,])/nrow(risk_ncdb)*100, 2)` %) died. 
In NCDB, the overall crude c-index values were lowest for D'Amico, NICE and EAU at 
`r round(min(unlist(ncdb_c_index$concord_crude)), 3)`, while the highest at 
`r round(max(unlist(ncdb_c_index$concord_crude)), 3)` was observed in GCP. In 
those who recieved a prostatectomy, the lowest crude concordance index was for 
AUA and the lowest adjusted concordance index was for the continuous CAPRA 
score, while the highest crude concordance was in NICE and EAU and the highest
adjusted concordance was in GUROC. 



**Discussion**

* PCa-specific mortality only avilable for SEER (not available for NCDB) - treat as sensitivity analysis 
* in TCGA - risk tools correlated with European ancestry estimates?  

**Conclusion**

\newpage

**Acknowledgements**

The authors would like to thank the participants of SEER and NCDB. 

**Author Contributions**

TAG designed and directed research; JHC and TAG acquired the data; JHC and TAG
analyzed and interpreted the data; JHC and TAG wrote the paper. 

**Additional Information**

**Funding:**

**Conflict of Interest:** The authors declare that they have no conflict of interest. 

**Informed Consent:** Informed consent was obtained...

**Data Availability:** The datasets analysed during the current study are 
available from SEER and NCDB with an approved application. 

\newpage

**Tables**

```{r table_ones}
library(tableone)
seer_vars <- c("AGE_DX", "YEAR_DX", "psa", "tstage", "gleason", 
               "damico", "nice", "eau", "GUROC", "AUA", "AUAi", 
               "NCCN", "CPG", "capra_score",
               "SRV_TIME_MON", "os")
table_one_seer <- CreateTableOne(vars = seer_vars, data = risk_seer)
kableone(table_one_seer)


ncdb_vars <- c("AGE", "YEAR_OF_DIAGNOSIS", "psa", "tstage", "gleason", 
               "damico", "nice", "eau", "GUROC", "AUA", "AUAi", 
               "NCCN", "CPG", "capra_score",
               "DX_LASTCONTACT_DEATH_MONTHS", "os")
table_one_ncdb <- CreateTableOne(vars = ncdb_vars, data = risk_ncdb)
kableone(table_one_ncdb)

```

\newpage

**Figures**

```{r bar_proportions}
library(viridis)

risk_seer %>% 
  select(PUBCSNUM, damico, nice, eau, GUROC, AUA, AUAi, NCCN, CPG) %>%
  gather("classifier", "value", -PUBCSNUM) %>%
  drop_na(value) %>%
  group_by(classifier) %>%
  mutate(total_number = n_distinct(PUBCSNUM)) %>%
  group_by(classifier, value) %>%  
  mutate(number = n_distinct(PUBCSNUM),
         perc = (n_distinct(PUBCSNUM)/total_number)*100) %>%
  distinct(classifier, value, .keep_all = TRUE) %>%
  ungroup() %>%
  mutate(value = factor(value, levels = c("Very High", "High", "Intermediate Unfavorable",
                                          "Intermediate", "Intermediate Favorable",
                                          "Low"))) %>%
  ggplot(aes(x = classifier, y = perc, fill = value)) + 
  geom_col() + 
  scale_fill_viridis_d(option = "C", name = "") + 
  labs(x = "Risk Classifier", y = "Percent (%)", 
       title = "SEER") + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        legend.position="bottom")

risk_ncdb %>% 
  select(PUF_CASE_ID, damico, nice, eau, GUROC, AUA, AUAi, NCCN, CPG) %>%
  gather("classifier", "value", -PUF_CASE_ID) %>%
  drop_na(value) %>%
  group_by(classifier) %>%
  mutate(total_number = n_distinct(PUF_CASE_ID)) %>%
  group_by(classifier, value) %>%  
  mutate(number = n_distinct(PUF_CASE_ID),
         perc = (n_distinct(PUF_CASE_ID)/total_number)*100) %>%
  distinct(classifier, value, .keep_all = TRUE) %>%
  ungroup() %>%
  mutate(value = factor(value, levels = c("Very High", "High", "Intermediate Unfavorable",
                                          "Intermediate", "Intermediate Favorable",
                                          "Low"))) %>%
  ggplot(aes(x = classifier, y = perc, fill = value)) + 
  geom_col() + 
  scale_fill_viridis_d(option = "C", name = "") + 
  labs(x = "Risk Classifier", y = "Percent (%)", 
       title = "NCDB") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        legend.position="bottom")

```

\newpage

**References**