---
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

library(drake)
library(tidyverse)

```

**Comparing prostate risk stratification tools**

Jordan H. Creed
Christelle 
Travis A. Gerke 

Department of Cancer Epidemiology, H. Lee Moffitt Cancer Center & Research Institute, 12902 USF Magnolia Drive, Tampa, FL, 33612, USA

Key Words - prostate; risk-startification

Corresponding author:
Travis A. Gerke, ScD
Department of Cancer Epidemiology
H. Lee Moffitt Cancer Center and Research Institute 
Tampa, FL 33612-9416
Phone: 
Fax:
Email: travis.gerke@moffitt.org

\newpage

**Abstract**

**Background:**

**Methods:**

**Results:**

**Conclusions:**

\newpage

**Introduction**

Pretreatment risk stratification tools are used to predict prostate cancer 
outcomes in men. The D'Amico risk startification system was originally proposed
in ...

We used Surveillance, Epidemiology, and End Results (SEER) and National Cancer
Database (NCDB) prostate cancer databases to compare the performance of the 
following pretreatment risk stratification tools: D'Amico, NICE, EAU and CAPRA. 

**Paticipants and Methods**

*SEER*

```{r seer_import}

loadd(seer_raw)

```

SEER is a population-based database developed by the National Cancer Institute,
US Center for Disease Control and Prevention and regional and state cancer 
registeries and includes cancer cases from 18 states [@doll2018]. 

*NCDB*

NCDB is the result of the joint efforts of the American College of Surgeons and
the American Cancer Society and covers almost 70% of incident cancer cases
[@boffa2017]. 

**Statistical Analysis**

Risk classification was calculated for each risk stratification tool. For 
participants with T stage greater or lesser than those used in the original 
cutpoints, the nearest original point was used. Risk classifications were
determined using the values available. Cox proportional hazard models
(HR) and 95% confidence intervals for overall survival were conducted for each
risk tool. In addition to crude model, models were adjusted for age (continuous),
year of diagnosis (continuous), Charlson Comorbidity Index (not available for 
SEER), marital status, education level (not available for SEER) and primary treatment.  

All analyses were performed using R (version 3.5.0). 

**Results**

*SEER*

A total of `r length(unique(seer_raw$PUBCSNUM))` men are included in the 
prostate cancer data for SEER. 

*NCDB*

**Discussion**

**Conclusion**

\newpage

**Acknowledgements**

The authors would like to thank the participants of SEER and NCDB. 

**Author Contributions**

TAG designed and directed research; JHC and TAG acquired the data; JHC and TAG
analyzed and interpreted the data; JHC and TAG wrote the paper. 

**Additional Information**

**Funding:**

**Conflict of Interest:** The authors declare that they have no conflict of interest. 

**Informed Consent:** Informed consent was obtained...

**Data Availability:** The datasets analysed during the current study are 
available from SEER and NCDB with an approved application. 

\newpage

**Tables**

```{r table_ones}
library(tableone)
```

\newpage

**Figures**

\newpage

**References**